---
title: 'Exploratory Data Analysis: WinRed'
author: "Seo-young Silvia Kim"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: '3'
    df_print: paged
  html_notebook:
    toc: yes
    toc_float: yes
    toc_depth: 3
    theme: united
    highlight: tango
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
categories <- c("winred")
source(here::here("R", "01_data_import.R"))
```

# Any empty URLs? 

```{r}
df_raw %>% filter(is.na(url)) %>% nrow()
```

```{r}
df_raw %>% filter(url == "") %>% nrow()
```

Yes, and 3 with completely empty data (very minor candidates) e.g., [https://ballotpedia.org/A._Wayne_Johnson](https://ballotpedia.org/A._Wayne_Johnson)

```{r}
setdiff(
  df_raw %>% filter(is.na(url)) %>% .$name, 
  df_raw %>% filter(!is.na(url)) %>% .$name
)
```

# Does WinRed directory contain changing URLs or are they static? 

```{r}
temp <- df_raw %>%
  group_by(name, race, url) %>%
  slice(1) %>%
  ungroup() %>%
  group_by(name, race) %>%
  filter(n() > 1) %>%
  arrange(name, race, url) %>%
  select(name, url, date, everything())
temp
```

`r round(length(unique(temp$name)) / length(unique(df_raw$name)) * 100, digits = 1)`% (`r length(unique(temp$name))` out of `r length(unique(df_raw$name))` entities) change URL link at some point in time. But most of these seem to be because of addition of "?sc=winred-directory". 

```{r}
temp <- df_raw %>%
  mutate(url = gsub("\\?sc=winred-directory", "", url)) %>%
  group_by(name, race, url) %>%
  slice(1) %>%
  ungroup() %>%
  group_by(name, race) %>%
  filter(n() > 1) %>%
  arrange(name, race, url) %>%
  select(name, url, date, everything())
temp
```

`r round(length(unique(temp$name)) / length(unique(df_raw$name)) * 100, digits = 1)`% (`r length(unique(temp$name))` out of `r length(unique(df_raw$name))` entities) change URL link at some point in time.

# Does any WinRed entity have multiple URLs in the same day?

Oh yikes, yes e.g., Michelle Steel, Young Kim, Lindsey Graham

```{r}
df_raw %>%
  group_by(name, race, url, date) %>%
  slice(1) %>%
  ungroup() %>%
  group_by(name, race, date) %>%
  filter(n() > 1) %>%
  nrow()
```

# Any entities across multiple races?

Yes: couple errors, couple real deal

```{r}
temp <- df_raw %>%
  group_by(name, race) %>%
  slice(1) %>%
  ungroup() %>%
  group_by(name) %>%
  filter(n() > 1)
nrow(temp)
temp
```

* Alison Hayden CA-16 should be changed to CA-15
* Jimmy Rodriguez really ran for [both AZ and VT](https://ballotpedia.org/Jimmy_Rodriguez)
* Liz Marty May and Matthew Morris is just a matter of at-large districts (SD-1 ---> SD-AL, DE-NaN ---> DE-AL)
* Scott Perry should be PA-10 instead of PA-4 (previous held position)
* Vern Buchanan should be FL-16 instead of FL-13 (previous held position)

# Any amount changes within same URL?

```{r}
wide_date <- df_raw %>%
  # To kick out duplicates
  mutate(url = gsub("\\?sc=winred-directory", "", url)) %>%
  mutate(name = gsub(" for Congress", "", name)) %>%
  mutate(
    race = ifelse(name == "Alison Hayden" & race == "CA-16", "CA-15", race),
    race = ifelse(name == "Liz Marty May" & race == "SD-1", "SD-AL", race),
    race = ifelse(name == "Matthew Morris" & race == "DE-NaN", "DE-AL", race),
    race = ifelse(name == "Scott Perry" & race == "PA-4", "PA-10", race),
    race = ifelse(name == "Vern Buchanan" & race == "FL-13", "FL-16", race)
  ) %>%
  # filter(!(name == "Jimmy Rodriguez" & race == "AZ-8")) %>%
  filter(!is.na(url)) %>%
  pivot_wider(
    names_from = date,
    values_from = portfolio,
    names_prefix = "date_",
    values_fn = list
  ) %>%
  clean_names() %>%
  arrange(name, url)
```

Same entity, multiple URLs

```{r}
sum(duplicated(wide_date$name))
```

# Amount changes within URL: entities/portfolio

```{r}
temp <- cross2(
  wide_date %>% select(-name, -race, -url) %>% names(),
  seq(nrow(wide_date))
) %>%
  map(
    ~ {
      out <- list()
      out[[paste0("row_", str_pad(.x[[2]], width = 3, pad = "0"))]] <-
        wide_date[.x[[2]], .x[[1]], drop = TRUE]
      return(out)
    }
  )
```

```{r}
menu_list <- seq(nrow(wide_date)) %>%
  map(
    ~ unique(
      temp[
        seq(1, length(temp), by = (ncol(wide_date) - 3))[.x]:
          seq(
            ncol(wide_date) - 3, length(temp),
            by = (ncol(wide_date) - 3)
          )[.x]
      ]
    ) %>%
      bind_rows() %>%
      .[, 1, drop = TRUE]
  ) %>%
  map(~ Filter(Negate(is.null), .x)) %>%
  map(~ Filter(function(x) any(!is.na(x)), .x))
```

```{r}
menu_freq <- menu_list %>%
  map(length) %>%
  unlist()
```

`r length(which(menu_freq > 1))` cases have portfolio change within the same URL.

```{r}
wide_date[which(menu_freq > 1), ]
```

* `r length(which(menu_freq < 1))` cases have no amount associated with URL: e.g., a Thank You page after loss (not a real fundraiser)

```{r}
wide_date[which(menu_freq < 1), ] %>% select(url, everything())
```

# Example

Alan Swain: checked with screenshots

```{r}
swain <- wide_date[8, ] %>%
  select(-name, -race, -url) %>% 
  map(~ paste(unlist(.x), collapse = "-")) %>% 
  unlist() %>%
  .[. != ""]
```

Best set of code to extract portfolio:

```{r}
seq(unique(rleid(swain))) %>%
  set_names(., .) %>%
  map(
    ~ tibble(
      min = names(swain)[min(which(.x == rleid(swain)))],
      max = names(swain)[max(which(.x == rleid(swain)))],
      amount = wide_date[
        8, names(swain)[min(which(.x == rleid(swain)))], drop = TRUE
      ] %>% 
        unlist() %>%
        paste0(collapse = "-")
    )
  ) %>%
  bind_rows(.id = "seq") %>%
  mutate_at(vars("min", "max"), ~ ymd(gsub("date|_", "", .x)))
```
