---
title: 'Exploratory Data Analysis: Right.us'
author: "Seo-young Silvia Kim"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: '3'
    df_print: paged
  html_notebook:
    toc: yes
    toc_float: yes
    toc_depth: 3
    theme: united
    highlight: tango
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
categories <- c("rightus")
source(here::here("R", "01_data_import.R"))
```

# Any empty URLs?

```{r}
df_raw %>% filter(is.na(url)) %>% nrow()
```

```{r}
df_raw %>% filter(url == "") %>% nrow()
```

# Anedot vs. non-Anedot links?

The peculiar thing about Right.us data is that it contains non-Anedot links
as well (even WinRed/revv in some cases!)

```{r}
df_raw <- df_raw %>%
  mutate(anedot = grepl("anedot", url))

prop(df_raw, "anedot")
```

What about by entity? 

```{r}
prop(df_raw %>% group_by(name, race, anedot) %>% slice(1), "anedot")
```

For consistency with comparison to other platforms, use only Anedot fundraisers. But check for flags in the entire data!

```{r}
rightus <- df_raw %>% filter(grepl("anedot", url))
```

# Does Right.us directory contain changing URLs or are they static?

```{r}
temp <- df_raw %>%
  group_by(name, race, url) %>%
  slice(1) %>%
  ungroup() %>%
  group_by(name, race) %>%
  filter(n() > 1) %>%
  arrange(name, race, url) %>%
  select(name, url, date, everything())
temp
```

`r round(length(unique(temp$name)) / length(unique(rightus$name)) * 100, digits = 1)`% (`r length(unique(temp$name))` out of `r length(unique(rightus$name))` entities) change URL link at some point in time. 

Note that it is *not* a simple addition of "/rightus/" in the middle of the URL. For example, McMorris (US House 2020 (Incumbent), Washington District 5):

```{r}
df_raw %>%
  filter(name == "Cathy McMorris Rodgers") %>%
  group_by(portfolio, url) %>%
  slice(1) %>%
  select(portfolio, url, date) %>%
  arrange(date, portfolio)
```

So this is not like the WinRed case where the only difference was the addition of "?sc=winred-directory".

# Does any Right.us entity have multiple URLs in the same day?

```{r}
df_raw %>%
  group_by(name, race, url, date) %>%
  slice(1) %>%
  ungroup() %>%
  group_by(name, race, date) %>%
  filter(n() > 1) %>%
  nrow()
```

# Any entities across multiple races?

Change of races or candidate status (e.g., primary ---> general). For example,

* Lt. Governor 2020 (Primary), North Carolina
* Lt. Governor 2020 (GOP Nominee), North Carolina

Let's not differentiate by candidate status. If same name, URL, and amount, the following should be treated equally.

Also, don't differentiate by race year, if data is collected in 2020. For example, the following:

* State House 2019 (Incumbent), Virginia District 33
* State House 2021 (Incumbent), Virginia District 33

Restore race data later if necessary

```{r}
temp <- df_raw %>%
  group_by(name, race) %>%
  slice(1) %>%
  ungroup() %>%
  group_by(name) %>%
  filter(n() > 1)
nrow(temp)
temp
```

What if we exclude that?

```{r}
df_raw <- df_raw %>%
  mutate(
    race_orig = race,
    race = str_remove_all(
      race, c(seq(2018, 2026), "\\(.*?\\)|,") %>% paste0(collapse = "|")
    ),
    race = trimws(str_replace_all(race, "\\s{2,}", " "))
  )
df_raw %>% sample_n(10)
```

```{r}
temp <- df_raw %>%
  group_by(name, race) %>%
  slice(1) %>%
  ungroup() %>%
  group_by(name) %>%
  filter(n() > 1) %>%
  arrange(name, race, url) %>%
  select(name, race, date, everything())
temp
```

`r length(unique(group_indices(temp)))` different people changed races. Some are legit changes e.g., from 2018 House run to a 2020 Senate run. Some are mistakes.

* [Art Robinson](https://ballotpedia.org/Art_Robinson) changed from OR-04 to State Senate 2 (O)
* [Ben Bullock](https://ballotpedia.org/Ben_Bullock) (O)
* [Bill Reineke](https://ballotpedia.org/Bill_Reineke) (O)
* [Brandon Beach](https://ballotpedia.org/Brandon_Beach) (O)
* [Brenda Dietrich](https://ballotpedia.org/Brenda_Dietrich) (O)
* [Brett Kokinadis](https://ballotpedia.org/Brett_Kokinadis) no record of having ran for US House New Mexico District 3 (X): "US House New Mexico District 3" ---> "US House New Mexico District 1"
* [Butch Miller](https://ballotpedia.org/Butch_Miller) no record of having ran for State Senate Georgia District 50 (X): "State Senate Georgia District 50" ---> "State Senate Georgia District 49"
* [Chris Gildon](https://ballotpedia.org/Chris_Gildon) (O)
* [Dan Hemmert](https://ballotpedia.org/Daniel_Hemmert) (O)
* [Danny Malouf](https://ballotpedia.org/Danny_Malouf) (O) but not the 17th district but the 14th; so fix that (X) so "US House Illinois District 17" ---> "US House Illinois District 14"
* [Darren Bailey](https://ballotpedia.org/Darren_Bailey) (O)
* [David Richter](https://ballotpedia.org/David_Richter_(New_Jersey)) (O)
* [Doug Collins](https://ballotpedia.org/Doug_Collins) (O)
* [Doug Smith](https://ballotpedia.org/Doug_Smith_(New_York)) at New York and [Doug Smith](https://ballotpedia.org/Doug_Smith_(West_Virginia)) at West Virginia (O)
* [Dr. Roger Marshall](https://ballotpedia.org/Roger_Marshall) (O)
* [Dusty Johnson](https://ballotpedia.org/Dusty_Johnson) is an at-large district mistake case (X) so "US House South Dakota District 1" ---> "US House South Dakota District AL"
* [Eric Nelson](https://ballotpedia.org/Eric_Nelson_(West_Virginia)) (O) Also, there is a Pennsylvania Eric Nelson, a separate entity.
* [George Phillips](https://ballotpedia.org/George_Phillips_(Ohio)) for Ohio and a [George Phillips](https://ballotpedia.org/George_Phillips_(New_York)) for New York. Both legit. (O)
* [J.R. Claeys](https://ballotpedia.org/J.R._Claeys) (O)
* [James Burchett](https://ballotpedia.org/James_Burchett) no record of having run for 178 (X)  so "State House Georgia District 178" ---> "State House Georgia District 176"
* [James P. Bradley](https://ballotpedia.org/James_P._Bradley_(California)) no record of having run for 28 (X) "US House California District 28" ---. "US House California District 33"
* [Jerry Carter](https://ballotpedia.org/Jerry_Carter_(North_Carolina)) at North Carolina and [Jerry Carter](https://ballotpedia.org/Jerry_Carter_(South_Carolina)) at South Carolina (O)
* [Jim Butler](https://ballotpedia.org/Jim_Butler_(West_Virginia)) at West Virginia and [Jim Butler](https://ballotpedia.org/James_Butler_(Ohio)) at Ohio (O)
* [Jim Marter](https://ballotpedia.org/James_Marter) no record of having run for 17 (X) so "US House Illinois District 17" ---> "US House Illinois District 14"
* [Joe Profit](https://ballotpedia.org/Joe_Profit) no record of having run for 7 (X): so "US House Georgia District 7" ---> "US House Georgia District 6"
* [John Clark](https://ballotpedia.org/John_Clark_(Tennessee)) at Tennessee and [John Clark](https://ballotpedia.org/John_Clark_(New_Mexico)) at New Mexico (O) 
* [John Kennedy](https://ballotpedia.org/John_Neely_Kennedy) in Senate and [John Kennedy](https://ballotpedia.org/John_F._Kennedy_(Georgia)) in Georgia (O)
* [John LaHood](https://ballotpedia.org/John_LaHood) no record of having run for 172 (X): so "State House Georgia District 172" ---> "State House Georgia District 175"
* [John Wilkinson](https://ballotpedia.org/John_Wilkinson) (X) Yes, ran for House 9th, but not State Senate 49th but 50th. So "State Senate Georgia District 49" ---> "State Senate Georgia District 50" (X)
* [Jon Burns](https://ballotpedia.org/Jon_G._Burns) no record of having run for 158 (X): so "State House Georgia District 158" ---> "State House Georgia District 159"
* [Justine Wadsack](https://ballotpedia.org/Justine_Wadsack) (O)
* [Kathie Hess Crouse](https://ballotpedia.org/Kathie_Hess_Crouse) did run for district 13 but this was in 2016; rather, a mistake (X): so "State Senate West Virginia District 13" ---> "State Senate West Virginia District 8"
* [Kelly Townsend](https://ballotpedia.org/Kelly_Townsend) (O)
* [Kim Thatcher](https://ballotpedia.org/Kim_Thatcher) (O)
* [Lance Harris](https://ballotpedia.org/Lance_Harris) (O)
* [Larry Liston](https://ballotpedia.org/Larry_Liston) (O)
* [Lee Murphy](https://ballotpedia.org/Lee_Murphy_(Delaware)) is an at-large district mistake case (X) so "US House Delaware District 1" ---> "US House Delaware District At-Large"
* [Lisa Barnes](https://ballotpedia.org/Lisa_Barnes) (O)
* [Louis Blessing](https://ballotpedia.org/Louis_W._Blessing,_III) (O)
* [Lynn Findley](https://ballotpedia.org/Lynn_Findley) (O)
* [Marjorie Greene](https://ballotpedia.org/Marjorie_Taylor_Greene) did for a while run for the 6th: see AJC article [here](https://www.ajc.com/blog/politics/another-republican-drops-out-georgia-6th-giving-handel-clear-shot-mcbath/tX1rufaXA8Fi2jOlcYoUoL/) (O)
* [Mark Johnson](https://ballotpedia.org/Mark_Johnson_(Arkansas)) for Arkansas and [Mark Johnson](https://ballotpedia.org/Mark_Johnson_(Minnesota)) for Minnesota (O)
* [Mark Romanchuk](https://ballotpedia.org/Mark_Romanchuk) (O)
* [Mark Walker](https://ballotpedia.org/Mark_Walker_(Maine)) for Maine and a [Mark Walker](https://ballotpedia.org/Mark_Walker_(North_Carolina)) for North Carolina. Both legit. (O)
* [Matt Gurtler](https://ballotpedia.org/Matt_Gurtler) (O)
* [Mike LiPetri](https://ballotpedia.org/Michael_LiPetri) (O)
* [Mike McKell](https://ballotpedia.org/Mike_McKell) (O)
* [Nancy Barto](https://ballotpedia.org/Nancy_K._Barto) (O)
* [Nick Freitas](https://ballotpedia.org/Nick_Freitas) (O)
* [Niraj Antani](https://ballotpedia.org/Niraj_Antani) (O)
* [Renee Erickson](https://ballotpedia.org/Renee_Erickson) (O)
* [Rosemary Becchi](https://ballotpedia.org/Rosemary_Becchi) (O)
* [Sandy Smith](https://ballotpedia.org/Sandy_Smith_(North_Carolina)): senate run not documented on Ballotpedia, but... uh... merch documented at [Amazon](https://www.amazon.com/Sandy-Smith-Senate-2020-Carolina/dp/B07T32SM44?customId=B07537HQXD&th=1) and [FEC statement of candidacy](https://docquery.fec.gov/cgi-bin/forms/S0NC00293/1318647/) (O)
* [Scott Fitzgerald](https://ballotpedia.org/Scott_Fitzgerald) (O)
* [Scott Taylor](https://ballotpedia.org/Scott_Taylor_(Virginia)) (O)
* [Sean Feucht](https://ballotpedia.org/Sean_Feucht) no record of having ran for the 1st district (X): so "US House California District 1" ---> "US House California District 3"
* [Steven Jarvis](https://ballotpedia.org/Steven_Jarvis_(North_Carolina)) (O) ---> prior office
* [Tom Killion](https://ballotpedia.org/Thomas_Killion) no record of having ran for district 7 (X): so "State Senate Pennsylvania District 7" ---> "State Senate Pennsylvania District 9"
* [Tom Phillips](https://ballotpedia.org/Tom_Phillips_(Kansas)) Kansas and [Tom Phillips](https://ballotpedia.org/Tom_Phillips_(New_Jersey)) New Jersey (O)
* [Trey Stewart](https://ballotpedia.org/Harold_Stewart) (O) ---> prior office
* [Vennia Francois](https://ballotpedia.org/Vennia_Francois) (O)
* [Warren Petersen](https://ballotpedia.org/Warren_Petersen) (O)
* [Willie Dove](https://ballotpedia.org/Willie_Dove) (O)

```{r}
df_raw <- df_raw %>%
  mutate(
    # Will deal with at-large districts separately
    race = case_when(
      name == "Brett Kokinadis" & race == "US House New Mexico District 3" ~ 
        "US House New Mexico District 1",
      name == "Butch Miller" & race == "State Senate Georgia District 50" ~ 
        "State Senate Georgia District 49",
      name == "Danny Malouf" & race == "US House Illinois District 17" ~ 
        "US House Illinois District 14",
      name == "James Burchett" & race == "State House Georgia District 178" ~ 
        "State House Georgia District 176",
      name == "James P. Bradley" & race == "US House California District 28" ~ 
        "US House California District 33",
      name == "Jim Marter" & race == "US House Illinois District 17" ~ 
        "US House Illinois District 14",
      name == "Joe Profit" & race == "US House Georgia District 7" ~ 
        "US House Georgia District 6",
      name == "John LaHood" & race == "State House Georgia District 172" ~ 
        "State House Georgia District 175",
      name == "John Wilkinson" & race == "State Senate Georgia District 49" ~
        "State Senate Georgia District 50",
      name == "Jon Burns" & race == "State House Georgia District 158" ~ 
        "State House Georgia District 159",
      name == "Kathie Hess Crouse" & 
        race == "State Senate West Virginia District 13" ~ 
        "State Senate West Virginia District 8",
      name == "Sean Feucht" & race == "US House California District 1" ~
        "US House California District 3",
      name == "Tom Killion" & race == "State Senate Pennsylvania District 7" ~
        "State Senate Pennsylvania District 9",
      TRUE ~ race
    )
  )
```

```{r}
assert_that(
  df_raw %>%
  group_by(name, race) %>%
  slice(1) %>%
  ungroup() %>%
  group_by(name) %>%
  filter(n() > 1) %>%
  nrow() <
    nrow(temp)
)
```

# At-large districts need additional treatment

As seen from the Dusty Johnson and Lee Murphy record, the at-large district expression is not uniform.

```{r}
df_raw %>%
  filter(
    grepl(
      state_df %>% filter(al_2010 == 1) %>% .$stname %>%
        paste0(" ", ., " ", collapse = "|"),
      race
    ) & 
      grepl("US House", race) & grepl("District 1$", race)
  ) %>%
  mutate(
    race = case_when(
      grepl("US House ", race) & 
        grepl(
          state_df %>% filter(al_2010 == 1) %>% .$stname %>%
            paste0(" ", ., " ", collapse = "|"),
          race
        ) ~ 
        gsub("District 1$", "District AL", race),
      TRUE ~ race
    )
  )
```

# Change of names

Sometimes, candidate name if changed e.g., Christine Bish ---> Chris Bish. There is no need to treat these as different instances. Need to be manually dealt because sometimes there are URLs with multiple entities. For example, [this link](https://secure.anedot.com/simonsen-and-mcclellan-for-south-jersey/c502f7dda9b7af22160a1). Also one wrong URL.

```{r}
df_raw %>%
  select(name, race, url) %>%
  group_by(url, name) %>%
  slice(1) %>%
  arrange(url) %>%
  group_by(url) %>%
  filter(first(name) != last(name))
```
```{r}
df_raw <- df_raw %>%
  mutate(
    name = case_when(
      name == "Bob Cupp" & race == "State House Ohio District 4" &
        grepl("http://bobcupp.org/donate/", url) ~ "Robert Cupp",
      name == "Christine Bish" & race == "US House, California District 6" &
        grepl("bish-for-congress/c4ebd2fa51523d54a7dc9", url) ~ "Chris Bish",
      name == "Bret Guthrie" & race == "US House, Kentucky District 2" &
        grepl("brett-guthrie-for-congress/c2f680721e268126f6a08", url) ~
      "Brett Guthrie",
      name == "Daniel Crenshaw" & race == "US House, Texas District 2" &
        grepl("crenshaw/c89f699a7ac209c4a0d2f", url) ~ "Dan Crenshaw",
      name == "Ronda Baldwin-Kennedy" &
        race == "US House, California District 26" &
        grepl("ronda-kennedy-for-congress/c1c126c031f520d67376e", url) ~
      "Ronda Kennedy",
      name == "Virginia House GOP" & race == "Caucus Committee Virginia" &
        grepl("vahousegop/c9f894490d6e3ffb63ea7", url) ~
      "Virginia House Republican Campaign Committee",
      TRUE ~ name
    ),
    url = case_when(
      grepl("new-hampshire-senate-republican-nominee-fund-2020", url) &
        name == "New Mexico Senate Republican Nominee Fund 2020" ~
      gsub("hampshire", "mexico", url),
      TRUE ~ url
    )
  )
```

